---
title: "ESPN FF"
output: html_document
---

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidymodels)
library(dplyr)
library(httr)
library(reticulate)
library(plotly)
library(RPostgres)
library(DBI)

# Setup Python
reticulate::py_config()
```

# Connect to GCP Postgres
```{r}
# Read in GCP connection info
gcp_info <- as.data.frame(read_yaml("gcp_info.yml"))
gcp_info <- gcp_info %>%
    mutate(
        dbname = as.character(dbname),
        host = as.character(host),
        port = as.character(port),
        user = as.character(user),
        password = as.character(password)
    )

# Connect to GCP
con <- dbConnect(
    RPostgres::Postgres(),
    dbname=gcp_info$dbname,
    host=gcp_info$host,
    port=gcp_info$port,
    user=gcp_info$user,
    password=gcp_info$password
)
```

# Load Python and Connect to League
```{python}
import pandas as pd
import numpy as np
from espn_api.football import League
import requests

# Connect to league
league = League(league_id=6699721, year=2020)
# Set current week
current_week = 11
```

# Python Functions
## fixTeams
```{python}
# Function to drop annoying team naming conventions
def fixTeams(teamName):
    newName = str(teamName).replace('Team(', '')
    newName = newName.replace(')','')
    return(newName)
```

## summarizeTeams
```{python}
# Function to convert teams to dataframe
def summarizeTeams(teams):
    # Vectors to hold info
    names = []
    wins = []
    losses = []
    standing = []
    points_for = []
    points_against = []
    points_diff = []
    
    # Loop through teams and extract info
    for team in teams:
        names.append(fixTeams(team.team_name))
        wins.append(team.wins)
        losses.append(team.losses)
        standing.append(team.standing)
        points_for.append(team.points_for)
        points_against.append(team.points_against)
        points_diff.append(team.points_for - team.points_against)
    
    # Create dataframe
    temp_ = {'name': names, 'wins': wins, 'losses': losses, 'standing': standing, 'points_for': points_for, 'points_against': points_against, 'points_diff': points_diff}
    teams = pd.DataFrame(temp_)
    return(teams)
```

## getPlayerPerformances
```{python}
# Function to get players performances
def getPlayerPerformances(current_week):
    # Vectors to hold data
    players = []
    projected = []
    points = []
    pts_diff = []
    perf_cat = []
    position = []
    weeks = []
    team = []
    # Iterate over weeks and compound data
    for i in range(1, current_week):
      # Get matchups from the week
      scores = league.box_scores(week=i)
      # Iterate over each game
      for game in scores:
        # Iterate over players on each team and get projected points
        for player in game.home_lineup:
          players.append(player.name)
          projected.append(player.projected_points)
          points.append(player.points)
          pts_diff.append(player.points - player.projected_points)
          perf_cat.append(catPerformance(player.projected_points, player.points))
          position.append(player.position)
          weeks.append(i)
          team.append(fixTeams(game.home_team))
        for player in game.away_lineup:
          players.append(player.name)
          projected.append(player.projected_points)
          points.append(player.points)
          pts_diff.append(player.points - player.projected_points)
          perf_cat.append(catPerformance(player.projected_points, player.points))
          position.append(player.position)
          weeks.append(i)
          team.append(fixTeams(game.away_team))

    # Create dataframe
    players_ = {'player': players, 'projected': projected, 'points': points, 'pts_diff': pts_diff, 'perf_cat': perf_cat, 'position': position, 'week': weeks, 'team': team}
    players = pd.DataFrame(players_) 
    
    # Return data
    return(players)
```

## catPerformance
```{python}
# Function to categorize performance
def catPerformance(proj_pts, pts):
    # Determine performance
    if proj_pts == 0:
        return("NO PROJECTION")
    elif (pts/proj_pts) > 1.5:
        return("OVERPERFORMED")
    elif (pts/proj_pts) < 0.5:
        return("UNDERPERFORMED")
    else:
        return("NEUTRAL")
```

# General R Code
## Define Python objects
```{python}
team_summary = summarizeTeams(league.teams)
performances = getPlayerPerformances(current_week)
```

## Convert to R
```{r}
# Convert to dataframes
team_summary <- as.data.frame(py$team_summary)
performances <- as.data.frame(py$performances)
```

# Create Aggregated Data Frames
## Grouped Performances Table
Grouping to a player level to get a player's averages for the season. Also counting how many weeks the particular play has been scored, as some players will be less than the season total due to pick-ups/drops in free agency.

### Aggregate Data
```{r}
grp_perf <- performances %>%
  group_by(player) %>%
  summarise(
    projected = mean(projected),
    points = mean(points),
    pts_diff = mean(pts_diff),
    weeks = n_distinct(week),
    ff_team = max(team),
    position = max(position)
  )
```

## FF Team Performances
Grouping to a positional level by week for each team in the league. Allows for an idea of how a team may be performing over time in particular positions.

### Aggregate Data
```{r}
ff_perf <- performances %>%
  group_by(team, week, position) %>%
  summarise(
    projected = mean(projected),
    points = mean(points),
    pts_diff = mean(pts_diff),
    underperformed = n_distinct(player[perf_cat == 'UNDERPERFORMED']),
    overperformed = n_distinct(player[perf_cat == 'OVERPERFORMED']),
    neutral = n_distinct(player[perf_cat == 'NEUTRAL'])
  )
```

## Position Performances
Aggregating to a position level by week to see potential trends in position groups over time.

### Aggregate Data
```{r}
position_grp <- performances %>%
  group_by(position, week) %>%
  summarise(
    players = n_distinct(player),
    projected = mean(projected),
    points = mean(points),
    pts_diff = mean(pts_diff),
    underperformed = n_distinct(player[perf_cat == 'UNDERPERFORMED']),
    overperformed = n_distinct(player[perf_cat == 'OVERPERFORMED']),
    neutral = n_distinct(player[perf_cat == 'NEUTRAL'])
  )
```

## Capture more Python Data
```{python}
games = league.box_scores(week = 11)
player_ids = []

# Extract players from current lineups
for game in games:
  for player in game.home_lineup:
    player_ids.append(player.playerId)
  for player in game.away_lineup:
    player_ids.append(player.playerId)

# Extract free agents
positions = ['QB','RB','WR','TE','K','D/ST']
fa_players = []

# Iterate over each position
for position in positions:
  # Get free agents
  fa = league.free_agents(week = 11, size = 100, position = position)
  # Loop over free agents
  for player in fa:
    fa_players.append(player.playerId)

# HANDLE PLAYERS ON CURRENT TEAMS
# Vectors to hold info
players = []
team = []
pos_rank = []
position = []

# Loop over players
for player in player_ids:
  temp = league.player_info(playerId = player)
  players.append(temp.name)
  team.append(temp.proTeam)
  pos_rank.append(temp.posRank)
  position.append(temp.position)
  
# Create dataframe
rostered_players_ = {'name': players, 'team': team, 'position_rank': pos_rank, 'position': position}
rostered_players = pd.DataFrame(rostered_players_) 
  
# HANDLE FREE AGENTS
# Two parts: 1) Get summary info 2) Get line-item info
# SUMMARY INFO
# Vectors to hold info
players = []
team = []
pos_rank = []
position = []

# Loop over free agents
for player in fa_players:
  temp = league.player_info(playerId = player)
  players.append(temp.name)
  team.append(temp.proTeam)
  pos_rank.append(temp.posRank)
  position.append(temp.position)
  
# Create dataframe
free_agents_ = {'name': players, 'team': team, 'position_rank': pos_rank, 'position': position}
free_agents = pd.DataFrame(free_agents_) 
  
# LINE-ITEM INFO
# Vectors to hold info
players = []
week = []
points = []

for player in fa_players:
  temp = league.player_info(playerId = player)
  # Grab stats data
  sts = temp.stats
  # Iterate over weeks
  for key in sts.keys():
    players.append(temp.name)
    week.append(str(key))
    points.append(sts[key]['points'])
    
# Create dataframe
fa_line_item_ = {'name': players, 'week': week, 'points': points}
fa_line_item = pd.DataFrame(fa_line_item_) 
```

## Convert to R
```{r}
# Convert to dataframes
rostered_players <- as.data.frame(py$rostered_players)
free_agents <- as.data.frame(py$free_agents)
fa_line_item <- as.data.frame(py$fa_line_item)
```

### Filter Data
```{r}
# Drop those with free agent lines outside of current week
fa_line_item <- fa_line_item %>%
  filter(week < py$current_week)
```

# Clean Data & Join Tables
Cleaning up final tables and joining in info before moving to Postgres

## Combine Rostered & Free Agents
```{r}
player_list <- rbind(rostered_players, free_agents)
```

## Player Line-Item Performances
```{r}
# Define columns to drop
drop_cols <- c("team")

# Subset rostered info
temp_player_list <- player_list %>%
  select(name, team)

# Drop FF tean from data
performances <- performances %>%
  select(-one_of(drop_cols))

# Join in team
performances <- performances %>%
  left_join(temp_player_list, by = c("player" = "name"))
```

## Player Aggregated Performances
```{r}
# Define columns to drop
drop_cols <- c("ff_team")

# Subset rostered info
temp_player_list <- player_list %>%
  select(name, team)

# Drop FF tean from data
grp_perf <- grp_perf %>%
  select(-one_of(drop_cols))

# Join in team
grp_perf <- grp_perf %>%
  left_join(temp_player_list, by = c("player" = "name"))
```

## Free Agents
```{r}
# Join free agent info with line item details
fa_line_item <- fa_line_item %>%
  left_join(free_agents, by = c("name"))
```

## Full Player List
```{r}
# Combine free agent list and performances list to create a full list of points performances for players. Data should contain all points scored, but not all points projected (free agents don't show week projections)
# Subset data
tmp_fa_line_item <- fa_line_item %>%
  select(name, week, points, team, position)
tmp_performances <- performances %>%
  select(player, points, week, team, position)

# Rename for binding purposes
names(tmp_performances) <- c("name", "points", "week", "team", "position")

# Bind
full_players <- rbind(tmp_fa_line_item, tmp_performances)
```

##################################################################
##################################################################
##################################################################

# Push Data to GCP Postgres
Send all the data into Postgres tables

## Add load dates to tables
```{r}
performances <- performances %>% mutate(load_d = lubridate::today())
full_players <- full_players %>% mutate(load_d = lubridate::today())
```

## Drop Tables
```{r}
# Dropping tables
dbRemoveTable(con, "performances")
dbRemoveTable(con, "grouped_performances")
dbRemoveTable(con, "ff_team_performances")
dbRemoveTable(con, "position_performances")
dbRemoveTable(con, "rostered_players")
dbRemoveTable(con, "free_agents")
dbRemoveTable(con, "fa_line_item")
dbRemoveTable(con, "full_players")
```

## Create Tables
```{r}
dbCreateTable(con, "performances", performances)
dbCreateTable(con, "grouped_performances", grp_perf)
dbCreateTable(con, "ff_team_performances", ff_perf)
dbCreateTable(con, "position_performances", position_grp)
dbCreateTable(con, "rostered_players", rostered_players)
dbCreateTable(con, "free_agents", free_agents)
dbCreateTable(con, "fa_line_item", fa_line_item)
dbCreateTable(con, "full_players", full_players)
```

## Write Tables
```{r}
dbWriteTable(con, "performances", performances, append = TRUE)
dbWriteTable(con, "grouped_performances", grp_perf, overwrite = TRUE)
dbWriteTable(con, "ff_team_performances", ff_perf, overwrite = TRUE)
dbWriteTable(con, "position_performances", position_grp, overwrite = TRUE)
dbWriteTable(con, "rostered_players", rostered_players, overwrite = TRUE)
dbWriteTable(con, "free_agents", free_agents, overwrite = TRUE)
dbWriteTable(con, "fa_line_item", fa_line_item, overwrite = TRUE)
dbWriteTable(con, "full_players", full_players, append = TRUE)
```
